SEER Medicaid

### 2023-09-07 Manuscript Analysis code 0 to 19.Rmd
added code to calculate p-values for stage analysis discontinuous vs. continuous comparison

# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```
# Import files sent by SEER Medicaid (IMS), get data into right columns, and assign column names
```{r, eval =FALSE}
# # Medicaid enrollment--imports fixed width columns and gives them names
# enroll <- read_fwf('medicaid.enroll.txt', 
#                fwf_cols(patient_id = 15, STATE_CD = 2, YEAR = 4, DOB_MATCH_SCORE_HI = 1, SEX_MATCH_SCORE_HI = 1,
#                        ID_MATCH_SCORE_HI=1, MATCH_SCORE_HI=2, MEDICAID_ELIG_CD1=1, MEDICAID_ELIG_CD2=1,
#                       MEDICAID_ELIG_CD3=1, MEDICAID_ELIG_CD4=1, MEDICAID_ELIG_CD5=1, MEDICAID_ELIG_CD6=1,
#                       MEDICAID_ELIG_CD7=1, MEDICAID_ELIG_CD8=1, MEDICAID_ELIG_CD9=1, MEDICAID_ELIG_CD10=1,
#                         MEDICAID_ELIG_CD11=1, MEDICAID_ELIG_CD12=1), 
#                skip = 0)

@@

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```
# Relevel to determine if there is a difference between continuous and discontinuous
```{r}
seer0to19_stage <- seer0to19_stage%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))
seer0to19_stage$mcont66.2 <- factor(seer0to19_stage$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

@@

>>>>

# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl)
```
# Functions used
```{r}
# function to calculate clustered SEs (from: https://stackoverflow.com/questions/73136246/clustered-standard-errors-stars-and-summary-statistics-in-modelsummary-for-mul)
mlogit.clust <- function(model,data,variable) {
  beta <- c(t(coef(model)))
  vcov <- vcov(model)
  k <- length(beta)
  n <- nrow(data)
  max_lev <- length(model$lev)
  xmat <- model.matrix(model)
  # u is deviance residuals times model.matrix
  u <- lapply(2:max_lev, function(x)
    residuals(model, type = "response")[, x] * xmat)
  u <- do.call(cbind, u)
  m <- dim(table(data[,variable]))
  u.clust <- matrix(NA, nrow = m, ncol = k)
  fc <- factor(data[,variable])
  for (i in 1:k) {
    u.clust[, i] <- tapply(u[, i], fc, sum)
  }
  cl.vcov <- vcov %*% ((m / (m - 1)) * t(u.clust) %*% (u.clust)) %*% vcov
  return(cl.vcov = cl.vcov)
}
```

# Import files sent by SEER Medicaid (IMS), get data into right columns, and assign column names
```{r, eval =FALSE}
# # Medicaid enrollment--imports fixed width columns and gives them names
# enroll <- read_fwf('medicaid.enroll.txt', 
#                fwf_cols(patient_id = 15, STATE_CD = 2, YEAR = 4, DOB_MATCH_SCORE_HI = 1, SEX_MATCH_SCORE_HI = 1,
#                        ID_MATCH_SCORE_HI=1, MATCH_SCORE_HI=2, MEDICAID_ELIG_CD1=1, MEDICAID_ELIG_CD2=1,
#                       MEDICAID_ELIG_CD3=1, MEDICAID_ELIG_CD4=1, MEDICAID_ELIG_CD5=1, MEDICAID_ELIG_CD6=1,
#                       MEDICAID_ELIG_CD7=1, MEDICAID_ELIG_CD8=1, MEDICAID_ELIG_CD9=1, MEDICAID_ELIG_CD10=1,
#                         MEDICAID_ELIG_CD11=1, MEDICAID_ELIG_CD12=1), 
#                skip = 0)

@@

rep_str2 = c('1: Regional:mcont66Continuous'='Continuous','1: Regional:mcont66Discontinuous'='Discontinuous','1: Regional:mcont66Other'='Other', '2: Distant:mcont66Continuous'='Continuous','2: Distant:mcont66Discontinuous'='Discontinuous','2: Distant:mcont66Other'='Other')
df16$Category<- str_replace_all(df16$Category, rep_str2)
df16 # this is the dataframe for summary results
```
# Table 2 with clustered standard errors used for confidence interval calculation
```{r}
Table2 <- rbind(df12, df16)
```

# Relevel to determine if there is a difference between continuous and discontinuous
```{r}
seer0to19_stage <- seer0to19_stage%>%
  mutate(mcont66.2 = case_when(mcont66 == "Continuous" ~ 0,
                               mcont66 == "Discontinuous" ~ 1,
                               mcont66 == "Other" ~ 2,
                               mcont66 == "Not enrolled" ~ 3))
seer0to19_stage$mcont66.2 <- factor(seer0to19_stage$mcont66.2, levels = c(0:3), labels = c("Continuous", "Discontinuous", "Other", "Not enrolled"))

@@

b <- c(t(coef(model2a)))
var <- mlogit.clust(model2a, seer0to19_stage, "state")
se <- sqrt(diag(var))
z <- b/se
p <- (1 - pnorm(abs(z), 0, 1))*2 # two-tailed z test, reported in manuscript
p
